### 2014-05-15 ApTranscriptome_TR.Rmd
Fixing analyses for manuscript

# knitr options
opts_chunk$set(cache=TRUE)
options(width = 80)

@@

@@

```{r sailfish_quant, eval=TRUE, echo=TRUE, cache=TRUE}
# directory containing trimmed reads
readdir <- "data/ind_files/" 
# sailfish index directory
sfindex <- "results/trinity-full/sailfish-index-Trinity-cap3-uclust"
# sailfish expression directory
sfexpressionroot <- "results/trinity-full/sailfish-expression-Trinity-cap3-uclust/"

# list of reads in each of four trimmed classes
readlist <- list.files(readdir)
(paired.left <- readlist[grep(".\\.paired.left.fastq$", readlist)])
(paired.right <- readlist[grep("\\.paired.right.fastq$", readlist)])
(unpaired.left <- readlist[grep("unpaired.left.fastq$", readlist)])
(unpaired.right <- readlist[grep("unpaired.right.fastq$", readlist)])

@@

A22.TPM[val==3,val:=3.5]
A22.TPM[val==10,val:=10.5]
A22.TPM[val==17,val:=17.5]
A22.TPM[val==24,val:=24.5]
A22.TPM[val==31,val:=31.5]
A22.TPM[val==38,val:=38.5]
head(A22.TPM)
str(A22.TPM)

# Repeat for Ar
Ar.TPM <- rbind(Ar_0_quant, Ar_3_quant, Ar_7_quant, Ar_10_quant, Ar_14_quant, Ar_17_quant, Ar_21_quant, Ar_24_quant, Ar_28_quant, Ar_31_quant, Ar_35_quant, Ar_38_quant)
Ar.TPM <- data.table(Ar.TPM)
setkey(Ar.TPM, Transcript, val)
Ar.TPM[val==3,val:=3.5]
Ar.TPM[val==10,val:=10.5]
Ar.TPM[val==17,val:=17.5]
Ar.TPM[val==24,val:=24.5]
Ar.TPM[val==31,val:=31.5]
Ar.TPM[val==38,val:=38.5]
head(Ar.TPM)
str(Ar.TPM)

tables()
```

@@

cortable <- cbind(temp, cors)
```

```{r exp_correlations_table, echo=FALSE, results='asis'}
pandoc.table(cortable, style="rmarkdown", caption = "correlations between colonies at each temperature treatment")
```

@@

```{r combined_TPM, echo=TRUE, eval=TRUE, cache=TRUE}
A22.TPM[,colony:="A22"]
Ar.TPM[,colony:="Ar"]
TPM.dt <- rbind(A22.TPM, Ar.TPM)
TPM.dt$colony <- as.factor(TPM.dt$colony)
str(TPM.dt)

@@

@@

# Pearson Chi-square test
chisq.test(Ap.type.table)

# Reorganize for plotting to show overlap among categories between colonies
type.table <- table(Ap.response.type[ , 'A22.type'], Ap.response.type[ , 'Ar.type'])

# reorder 
tt2 <- type.table[c(2,4,1,3,5), c(2,4,1,3,5)]

# plot
mosaicplot(tt2, xlab = "AcNC", ylab = "ApVT", main = "Mosaic plot of responsive transcripts")
```

@@

```{r temperature_response_table, echo=FALSE, results='asis'}
pandoc.table(tt2, style="rmarkdown", caption = "Number of transcripts with maximum expression at high, low, intermediate, both high and low (bimodal) temperatures or are not thermally-responsivefor each colony and their overlap.", split.tables = "Inf")
```

@@

### Plasticity versus constitutive expression

Selection may have acted in response to thermal stress such that some genes are constitutively activated in one colony while plastically-expressed in the other. I evaluate this hypothesis by comparing expression levels at the at optimum (19.5C) temperature between the two colonies for genes in that are either in the 'High' or 'Low' expression group in the other colony. Specifically, I expect that genes upregulated at high temperatures in *A22* are more highly expressed at 19.5C in *Ar*?

@@

The non-parametric test for both comparisions also finds greater constitutive expression in *Ar* than *A22*.

@@

# calculate differences in mean, slope, curvature and wiggle between colonies for each transcript
# offset = mean1 - mean2
varshape.out$offset <- varshape.out$A22.mean - varshape.out$Ar.mean
# to compare among transcripts that differ in scale of expression, standardize by overall mean for each transcript
varshape.out$s.offset <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.offset"] <- (varshape.out[i, "offset"] / mean(c(varshape.out[i, "A22.mean"], varshape.out[i, "Ar.mean"])))
}

# t-test
(t3 <- t.test(varshape.out$s.offset, alternative = "two.sided"))

# slope = slope1 - slope2
varshape.out$slope <- varshape.out$A22.slope - varshape.out$Ar.slope
# to compare among transcripts that differ in scale of expression, standardize by overall mean for each transcript
varshape.out$s.slope <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.slope"] <- (varshape.out[i, "slope"] / mean(c(varshape.out[i, "A22.slope"], varshape.out[i, "Ar.slope"])))
}
# for transcripts where both samples had slope=0, change s.slope from Inf to 0 as there is truly no difference in slope
varshape.out[which(varshape.out$Ar.slope == 0 & varshape.out$A22.slope == 0), "s.slope"] <- 0

# t-test
(t4 <- t.test(varshape.out$s.slope, alternative = "two.sided"))

# curvature = curvature1 - curvature2
varshape.out$curvature <- varshape.out$A22.curve - varshape.out$Ar.curve
# to compare among transcripts that differ in scale of expression, standardize by overall mean for each transcript
varshape.out$s.curvature <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.curvature"] <- (varshape.out[i, "curvature"] / mean(c(varshape.out[i, "A22.curve"], varshape.out[i, "Ar.curve"])))
}
# change 'Inf' values to 0
varshape.out[which(varshape.out$Ar.curve == 0 & varshape.out$A22.curve == 0), "s.curvature"] <- 0
# t-test
(t.test(varshape.out$s.curvature, alternative = "two.sided"))

# wiggle
varshape.out$wiggle <- varshape.out$A22.wiggle - varshape.out$Ar.wiggle
varshape.out$s.wiggle <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.wiggle"] <- (varshape.out[i, "wiggle"] / mean(c(varshape.out[i, "A22.wiggle"], varshape.out[i, "Ar.wiggle"])))
}
# change 'Inf' values to 0
varshape.out[which(varshape.out$Ar.wiggle == 0 & varshape.out$A22.wiggle == 0), "s.wiggle"] <- 0
# note that only 407 transcripts have any 'wiggle'
length(which(varshape.out$s.wiggle != 0))

# t-test
(t5 <- t.test(varshape.out$s.wiggle, alternative = "two.sided"))
```

@@

```{r partition_reaction_norm_variation}
# weighted mean-standardized values of each measure
mean(abs((varshape.out$s.offset)))
mean(abs((varshape.out$s.slope)))
mean(abs((varshape.out$s.curvature)))
mean(abs((varshape.out$s.wiggle)))

@@

mean(varshape.out$prop.wiggle)
quantile(varshape.out$prop.wiggle, probs = c(0.05, 0.5, 0.95))
```

@@

```{r responsive_TPM}
# extract TPM data for thermally-responsive transcripts
resp.TPM.dt.sub <- TPM.dt.sub[names(responsive.lms)]
setkey(resp.TPM.dt.sub, Transcript)
str(resp.TPM.dt.sub)
length(unique(resp.TPM.dt.sub$Transcript))
# scale transcripts so can compare
resp.TPM.dt.sub[,TPM.scaled:=scale(TPM), by = Transcript]
```

@@

intersect(A22.low.TPM.dt.sub, A22.high.TPM.dt.sub)
# Good - no overlap

@@

Genes with increased expression at *Low* temperatures are turned on at lower temperatures in *AcNC* than *ApVT*.

@@

### Identification of biological-functions associated with temperature responses

## Functional annotation

@@

```{r echo=FALSE, results='hide'}
responsive.lms.ann.type <- signif.transcripts[names(responsive.lms)]
# combine responsive.lms.ann with "type" information
responsive.lms.ann.type <- merge(responsive.lms.ann.type, Ap.response.type[,c("Transcript", "Ar.type", "Ar.type")], by = "Transcript", all = TRUE)
str(responsive.lms.ann.type)
write.csv(responsive.lms.ann.type, file = paste(resultsdir, "Ap_responsive_genes_", Sys.Date(), ".csv", sep = ""), row.names = FALSE)
```

## Gene set enrichment analysis ##

@@

Ap.BP.GOdata

@@

# table results
Ap.BP.ResTable <- GenTable(Ap.BP.GOdata, parentchild = Ap.BP.resultParentChild, topNodes = 137)
dim(Ap.BP.ResTable)
#pandoc.table(Ap.BP.ResTable)

# graph significant nodes

#pdf(paste(resultsdir, "Ap.BP_topGO_sig_nodes.pdf", sep=""))
#showSigOfNodes(Ap.BP.GOdata, score(Ap.BP.resultParentChild), firstSigNodes = 10, useInfo = 'all')
#dev.off()
```

@@

```{r gsea_high}
# Select transcripts
Ap.high <- Ap.response.type[which(Ap.response.type$Ar.type == "High" & Ap.response.type$A22.type == "High"), "Transcript"]
Ap.geneList.high <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.high) <- names(Ap.geneList)
Ap.geneList.high[(which(names(Ap.geneList.high) %in% Ap.high))] <- 1
# check correct number of values set to 1
table(Ap.geneList.high)

@@

```{r gsea_low}
# Select transcripts
Ap.low <- Ap.response.type[which(Ap.response.type$Ar.type == "Low" & Ap.response.type$A22.type == "Low"), "Transcript"]
Ap.geneList.low <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.low) <- names(Ap.geneList)
Ap.geneList.low[(which(names(Ap.geneList.low) %in% Ap.low))] <- 1
# check correct number of values set to 1
table(Ap.geneList.low)

@@

```{r gsea_bimodal}
Ap.bim <- Ap.response.type[which(Ap.response.type$A22.type == "Bimodal" & Ap.response.type$Ar.type == "Bimodal"), "Transcript"]
# create gene list, setting value to 1 for "bim" transcripts
Ap.geneList.bim <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.bim) <- names(Ap.geneList)
Ap.geneList.bim[(which(names(Ap.geneList.bim) %in% Ap.bim))] <- 1
# check correct number of values set to 1
table(Ap.geneList.bim)

@@

```{r gsea_intermediate}
Ap.int <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate" & Ap.response.type$Ar.type == "Intermediate"), "Transcript"]
# create gene list, setting value to 1 for "int" transcripts
Ap.geneList.int <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.int) <- names(Ap.geneList)
Ap.geneList.int[(which(names(Ap.geneList.int) %in% Ap.int))] <- 1
# check correct number of values set to 1
table(Ap.geneList.int)

@@

```{r gsea_high_A22}
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.high <- rep(0, length = length(Ap.geneList))
names(A22.geneList.high) <- names(Ap.geneList)
A22.geneList.high[(which(names(A22.geneList.high) %in% A22.high))] <- 1
# check correct number of values set to 1
table(A22.geneList.high)

@@

```{r gsea_high_Ar}
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.high <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.high) <- names(Ap.geneList)
Ar.geneList.high[(which(names(Ar.geneList.high) %in% Ar.high))] <- 1
# check correct number of values set to 1
table(Ar.geneList.high)

@@

```{r gsea_low_A22}
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.low <- rep(0, length = length(Ap.geneList))
names(A22.geneList.low) <- names(Ap.geneList)
A22.geneList.low[(which(names(A22.geneList.low) %in% A22.low))] <- 1
# check correct number of values set to 1
table(A22.geneList.low)

@@

```{r gsea_low_Ar}
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.low <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.low) <- names(Ap.geneList)
Ar.geneList.low[(which(names(Ar.geneList.low) %in% Ar.low))] <- 1
# check correct number of values set to 1
table(Ar.geneList.low)

@@

```{r gsea_bim_A22}
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.bim <- rep(0, length = length(Ap.geneList))
names(A22.geneList.bim) <- names(Ap.geneList)
A22.geneList.bim[(which(names(A22.geneList.bim) %in% A22.bim))] <- 1
# check correct number of values set to 1
table(A22.geneList.bim)

@@

```{r gsea_bim_Ar}
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.bim <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.bim) <- names(Ap.geneList)
Ar.geneList.bim[(which(names(Ar.geneList.bim) %in% Ar.bim))] <- 1
# check correct number of values set to 1
table(Ar.geneList.bim)

@@

```{r gsea_int_A22}
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.int <- rep(0, length = length(Ap.geneList))
names(A22.geneList.int) <- names(Ap.geneList)
A22.geneList.int[(which(names(A22.geneList.int) %in% A22.int))] <- 1
# check correct number of values set to 1
table(A22.geneList.int)

@@

```{r gsea_int_Ar}
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.int <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.int) <- names(Ap.geneList)
Ar.geneList.int[(which(names(Ar.geneList.int) %in% Ar.int))] <- 1
# check correct number of values set to 1
table(Ar.geneList.int)

@@

# Line plot, expression against temp, faceted by colony
trp1 <- ggplot(resp.TPM.dt.sub[high.stress.terms$Transcript], aes(x=val, y=TPM.scaled, group=Transcript)) + 
  geom_line() + 
  facet_grid(. ~ colony) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp1)

# Smooth plot, expression against temp, faceted by colony
trp2 <- ggplot(resp.TPM.dt.sub[high.stress.terms$Transcript], aes(x=val, y=TPM.scaled, group=Transcript)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  facet_grid(. ~ colony) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp2)
```

@@

# Line plot, expression against temp, faceted by colony
trp3 <- ggplot(resp.TPM.dt.sub[low.stress.terms$Transcript], aes(x=val, y=TPM.scaled, group=Transcript)) + 
  geom_line() + 
  facet_grid(. ~ colony) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp3)

# Smooth plot, expression against temp, faceted by colony
trp4 <- ggplot(resp.TPM.dt.sub[low.stress.terms$Transcript], aes(x=val, y=TPM.scaled, group=Transcript)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  facet_grid(. ~ colony) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp4)
```

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(tt2, Ap.df, w1, w2, Ap.T_on, file = "ms.RData")
```

>>>>

# knitr options
opts_chunk$set(cache=TRUE)

@@

To examine the species distribution of BLAST hits in the transcriptome assembly, I used the program [Krona](http://sourceforge.net/p/krona/home/krona/) `r citep("doi:10.1186/1471-2105-12-385")`. I ...

~~~
KRONA code
~~~

The interactive visualization is available [here]().

@@

```{r sailfish_quant, results = 'hide'}
# directory containing trimmed reads
readdir <- "data/ind_files/" 
# sailfish index directory
sfindex <- "results/trinity-full/sailfish-index-Trinity-cap3-uclust"
# sailfish expression directory
sfexpressionroot <- "results/trinity-full/sailfish-expression-Trinity-cap3-uclust/"

# list of reads in each of four trimmed classes
readlist <- list.files(readdir)
paired.left <- readlist[grep(".\\.paired.left.fastq$", readlist)]
paired.right <- readlist[grep("\\.paired.right.fastq$", readlist)]
unpaired.left <- readlist[grep("unpaired.left.fastq$", readlist)]
unpaired.right <- readlist[grep("unpaired.right.fastq$", readlist)]

@@

A22.TPM[val==3,val:=3.5]
A22.TPM[val==10,val:=10.5]
A22.TPM[val==17,val:=17.5]
A22.TPM[val==24,val:=24.5]
A22.TPM[val==31,val:=31.5]
A22.TPM[val==38,val:=38.5]
str(A22.TPM)

# Repeat for Ar
Ar.TPM <- rbind(Ar_0_quant, Ar_3_quant, Ar_7_quant, Ar_10_quant, Ar_14_quant, Ar_17_quant, Ar_21_quant, Ar_24_quant, Ar_28_quant, Ar_31_quant, Ar_35_quant, Ar_38_quant)
Ar.TPM <- data.table(Ar.TPM)
setkey(Ar.TPM, Transcript, val)
Ar.TPM[val==3,val:=3.5]
Ar.TPM[val==10,val:=10.5]
Ar.TPM[val==17,val:=17.5]
Ar.TPM[val==24,val:=24.5]
Ar.TPM[val==31,val:=31.5]
Ar.TPM[val==38,val:=38.5]
str(Ar.TPM)
```

@@

cortable <- cbind(Temperature = temp, Correlation = cors)
```

```{r exp_correlations_table, echo=FALSE, results='asis'}
pandoc.table(cortable, style="rmarkdown", caption = "Correlations between species for gene expression at temperature treatment")
```

@@

```{r combined_TPM, results = 'hide'}
A22.TPM[,colony:="A22"]
Ar.TPM[,colony:="Ar"]
TPM.dt <- rbind(A22.TPM, Ar.TPM)
TPM.dt$colony <- as.factor(TPM.dt$colony)
str(TPM.dt)

@@

DeconSeq identified contaminant genes. I remove this n

@@

# Pearson Chi-square test
chi1 <- chisq.test(Ap.type.table)
chi1

# Reorganize for plotting to show overlap among categories between colonies
type.table <- table(Ap.response.type[ , 'Ar.type'], Ap.response.type[ , 'A22.type'])
# reorder 
tt2 <- type.table[c(2,4,1,3,5), c(2,4,1,3,5)]

# plot
mosaicplot(t(tt2), ylab = "AcNC", xlab = "ApVT", main = "Mosaic plot of responsive transcripts")
```

@@

```{r temperature_response_table, echo=FALSE, results='asis'}
# add row and column totals
tt3 <- rbind(tt2, colSums(tt2))
tt3 <- cbind(tt3, rowSums(tt3))
colnames(tt3)[6] <- "Total"
rownames(tt3)[6] <- "Total"
# move names into table
tt3 <- rbind(rownames(tt3), tt3)
tt3 <- cbind(rownames(tt3), tt3)
# name col/row by species
rownames(tt3) <- c("AcNC", "", "", "", "", "", "")
colnames(tt3) <- c("ApVT", "", "", "", "", "", "")

pandoc.table(tt3, style="rmarkdown", caption = "Number of transcripts with maximum expression at high, low, intermediate, both high and low (bimodal) temperatures or are not thermally-responsivefor each colony and their overlap.", split.tables = "Inf")
```

@@

### Plasticity versus basal expression

Selection may have acted in response to thermal stress such that some genes are basaly activated in one colony while plastically-expressed in the other. I evaluate this hypothesis by comparing expression levels at the at optimum (19.5C) temperature between the two colonies for genes in that are either in the 'High' or 'Low' expression group in the other colony. Specifically, I expect that genes upregulated at high temperatures in *A22* are more highly expressed at 19.5C in *Ar*?

@@

The non-parametric test for both comparisions also finds greater expression in *Ar* than *A22* at the optimal temperature.

@@

t.mean <- t.test(log(varshape.out$A22.mean), log(varshape.out$Ar.mean))

# calculate differences in mean, slope, curvature and wiggle between colonies for each transcript
# offset = mean1 - mean2
varshape.out$offset <- varshape.out$A22.mean - varshape.out$Ar.mean

# t-test
(t.offset <- t.test(varshape.out$s.offset, alternative = "two.sided"))

# slope = slope1 - slope2
varshape.out$slope <- varshape.out$A22.slope - varshape.out$Ar.slope

# t-test
(t.slope <- t.test(varshape.out$s.slope, alternative = "two.sided"))

# curvature = curvature1 - curvature2
varshape.out$curvature <- varshape.out$A22.curve - varshape.out$Ar.curve
# to compare among transcripts that differ in scale of expression, standardize by overall mean for each transcript
varshape.out$s.curvature <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.curvature"] <- (varshape.out[i, "curvature"] / mean(c(varshape.out[i, "A22.curve"], varshape.out[i, "Ar.curve"])))
}
# change 'Inf' values to 0
varshape.out[which(varshape.out$Ar.curve == 0 & varshape.out$A22.curve == 0), "s.curvature"] <- 0
# t-test
(t.curvature <- t.test(varshape.out$s.curvature, alternative = "two.sided"))

# wiggle
varshape.out$wiggle <- varshape.out$A22.wiggle - varshape.out$Ar.wiggle
varshape.out$s.wiggle <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.wiggle"] <- (varshape.out[i, "wiggle"] / mean(c(varshape.out[i, "A22.wiggle"], varshape.out[i, "Ar.wiggle"])))
}
# change 'Inf' values to 0
varshape.out[which(varshape.out$Ar.wiggle == 0 & varshape.out$A22.wiggle == 0), "s.wiggle"] <- 0
# note small number of genes with 'wiggle'
length(which(varshape.out$s.wiggle != 0))

# t-test
(t.wiggle <- t.test(varshape.out$s.wiggle, alternative = "two.sided"))
```

@@

```{r partition_reaction_norm_variation}
# to compare among transcripts that differ in scale of expression, standardize by overall mean for each transcript
varshape.out$s.offset <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.offset"] <- (varshape.out[i, "offset"] / mean(c(varshape.out[i, "A22.mean"], varshape.out[i, "Ar.mean"])))
}
# to compare among transcripts that differ in scale of expression, standardize by overall mean for each transcript
varshape.out$s.slope <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.slope"] <- (varshape.out[i, "slope"] / mean(c(varshape.out[i, "A22.slope"], varshape.out[i, "Ar.slope"])))
}
# for transcripts where both samples had slope=0, change s.slope from Inf to 0 as there is truly no difference in slope
varshape.out[which(varshape.out$Ar.slope == 0 & varshape.out$A22.slope == 0), "s.slope"] <- 0

                                        # weighted mean-standardized values of each measure
mean(abs((varshape.out$s.offset)))
mean(abs((varshape.out$s.slope)))
mean(abs((varshape.out$s.curvature)))
mean(abs((varshape.out$s.wiggle)))

@@

mean(varshape.out$prop.wiggle)
qwiggle <- quantile(varshape.out$prop.wiggle, probs = c(0.05, 0.5, 0.95))
```

@@

```{r responsive_TPM}
# extract TPM data for thermally-responsive transcripts
resp.TPM.dt.sub <- TPM.dt.sub[names(responsive.lms)]
setkey(resp.TPM.dt.sub, Transcript)
str(resp.TPM.dt.sub)
length(unique(resp.TPM.dt.sub$Transcript))
# scale transcripts so can compare
resp.TPM.dt.sub[,TPM.scaled:=scale(TPM), by = Transcript]
# rename colonies
resp.TPM.dt.sub$colony2 <- ifelse(resp.TPM.dt.sub$colony == "A22", "ApVT", "AcNC")
```

@@

@@

Genes with increased expression at *High* temperatures are turned on at, on average, 1°C higher in *ApVT* than *AcNC*.

@@

## Gene set enrichment analysis for thermal-responsive genes

### Functional annotation

@@

```{r ann_responsive, echo=TRUE}
responsive.lms.ann.type <- signif.transcripts[names(responsive.lms)]
# combine responsive.lms.ann with "type" information
responsive.lms.ann.type <- merge(responsive.lms.ann.type, Ap.response.type[,c("Transcript", "A22.type", "Ar.type")], by = "Transcript", all = TRUE)
str(responsive.lms.ann.type)
write.csv(responsive.lms.ann.type, file = paste(resultsdir, "Ap_responsive_genes_", Sys.Date(), ".csv", sep = ""), row.names = FALSE)
```

@@

#Ap.BP.GOdata

@@

# table results
Ap.BP.ResTable <- GenTable(Ap.BP.GOdata, parentchild = Ap.BP.resultParentChild, topNodes = 10)
Ap.BP.ResTable
```

@@

```{r gsea_high, results = 'hide'}
# Select transcripts
Ap.high <- Ap.response.type[which(Ap.response.type$Ar.type == "High" & Ap.response.type$A22.type == "High"), "Transcript"]
Ap.geneList.high <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.high) <- names(Ap.geneList)
Ap.geneList.high[(which(names(Ap.geneList.high) %in% Ap.high))] <- 1
# check correct number of values set to 1
table(Ap.geneList.high)

@@

```{r gsea_low, results = 'hide'}
# Select transcripts
Ap.low <- Ap.response.type[which(Ap.response.type$Ar.type == "Low" & Ap.response.type$A22.type == "Low"), "Transcript"]
Ap.geneList.low <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.low) <- names(Ap.geneList)
Ap.geneList.low[(which(names(Ap.geneList.low) %in% Ap.low))] <- 1
# check correct number of values set to 1
table(Ap.geneList.low)

@@

```{r gsea_bimodal, results = 'hide'}
Ap.bim <- Ap.response.type[which(Ap.response.type$A22.type == "Bimodal" & Ap.response.type$Ar.type == "Bimodal"), "Transcript"]
# create gene list, setting value to 1 for "bim" transcripts
Ap.geneList.bim <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.bim) <- names(Ap.geneList)
Ap.geneList.bim[(which(names(Ap.geneList.bim) %in% Ap.bim))] <- 1
# check correct number of values set to 1
table(Ap.geneList.bim)

@@

```{r gsea_intermediate, results = 'hide'}
Ap.int <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate" & Ap.response.type$Ar.type == "Intermediate"), "Transcript"]
# create gene list, setting value to 1 for "int" transcripts
Ap.geneList.int <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.int) <- names(Ap.geneList)
Ap.geneList.int[(which(names(Ap.geneList.int) %in% Ap.int))] <- 1
# check correct number of values set to 1
table(Ap.geneList.int)

@@

```{r gsea_high_A22, results = 'hide'}
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.high <- rep(0, length = length(Ap.geneList))
names(A22.geneList.high) <- names(Ap.geneList)
A22.geneList.high[(which(names(A22.geneList.high) %in% A22.high))] <- 1
# check correct number of values set to 1
table(A22.geneList.high)

@@

```{r gsea_high_Ar, results = 'hide'}
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.high <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.high) <- names(Ap.geneList)
Ar.geneList.high[(which(names(Ar.geneList.high) %in% Ar.high))] <- 1
# check correct number of values set to 1
table(Ar.geneList.high)

@@

```{r gsea_low_A22, results = 'hide'}
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.low <- rep(0, length = length(Ap.geneList))
names(A22.geneList.low) <- names(Ap.geneList)
A22.geneList.low[(which(names(A22.geneList.low) %in% A22.low))] <- 1
# check correct number of values set to 1
table(A22.geneList.low)

@@

```{r gsea_low_Ar, results = 'hide'}
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.low <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.low) <- names(Ap.geneList)
Ar.geneList.low[(which(names(Ar.geneList.low) %in% Ar.low))] <- 1
# check correct number of values set to 1
table(Ar.geneList.low)

@@

```{r gsea_bim_A22, results = 'hide'}
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.bim <- rep(0, length = length(Ap.geneList))
names(A22.geneList.bim) <- names(Ap.geneList)
A22.geneList.bim[(which(names(A22.geneList.bim) %in% A22.bim))] <- 1
# check correct number of values set to 1
table(A22.geneList.bim)

@@

```{r gsea_bim_Ar, results = 'hide'}
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.bim <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.bim) <- names(Ap.geneList)
Ar.geneList.bim[(which(names(Ar.geneList.bim) %in% Ar.bim))] <- 1
# check correct number of values set to 1
table(Ar.geneList.bim)

@@

```{r gsea_int_A22, results = 'hide'}
# create gene list, setting value to 1 for "bim" transcripts
A22.geneList.int <- rep(0, length = length(Ap.geneList))
names(A22.geneList.int) <- names(Ap.geneList)
A22.geneList.int[(which(names(A22.geneList.int) %in% A22.int))] <- 1
# check correct number of values set to 1
table(A22.geneList.int)

@@

```{r gsea_int_Ar, results = 'hide'}
# create gene list, setting value to 1 for "bim" transcripts
Ar.geneList.int <- rep(0, length = length(Ap.geneList))
names(Ar.geneList.int) <- names(Ap.geneList)
Ar.geneList.int[(which(names(Ar.geneList.int) %in% Ar.int))] <- 1
# check correct number of values set to 1
table(Ar.geneList.int)

@@

# Line plot, expression against temp, faceted by colony
trp1 <- ggplot(resp.TPM.dt.sub[high.stress.terms$Transcript], aes(x=val, y=TPM.scaled, group=Transcript)) + 
  geom_line() + 
  facet_grid(. ~ colony2) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp1)

# Smooth plot, expression against temp, faceted by colony
trp2 <- ggplot(resp.TPM.dt.sub[high.stress.terms$Transcript], aes(x=val, y=TPM.scaled, group=Transcript)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  facet_grid(. ~ colony2) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp2)
```

@@

# Line plot, expression against temp, faceted by colony
trp3 <- ggplot(resp.TPM.dt.sub[low.stress.terms$Transcript], aes(x=val, y=TPM.scaled, group=Transcript)) + 
  geom_line() + 
  facet_grid(. ~ colony2) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp3)

# Smooth plot, expression against temp, faceted by colony
trp4 <- ggplot(resp.TPM.dt.sub[low.stress.terms$Transcript], aes(x=val, y=TPM.scaled, group=Transcript)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  facet_grid(. ~ colony2) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp4)
```

```{r ggplot_high}
trp_A22_high <- ggplot(resp.TPM.dt.sub[A22.high], aes(x=val, y=TPM.scaled, group=Transcript)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) + 
  facet_grid(. ~ colony2) + 
  scale_y_continuous(name="Expression (scaled)") +
  scale_x_continuous(name=expression(paste("Temperature ", degree, "C")))
print(trp_A22_high)

```

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(RxNpval, sigtable, chi1, resp.TPM.dt.sub, tt2, tt3, GO.stress.Xsq, GO0006950.responsive, Ap.df, w1, w2, Ap.T_on, Ap.BP.resultParentChild, Ap.gsea1, Ap.setdiff.gsea, resp.stress.responsive.count, t.offset, t.slope, t.curvature, t.wiggle, file = "ms.RData")
```
